---
title: "Processing M50 JTMS data"
author: "Dan Brennan"
date: "09 March 2017"
output: github_document
---

This note outlines the process and code used to read the M50 JTMS data into a database and output journey time values to be used for the calculation of performance indicators. 

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
```

The path to where the journey time data is defined and a connection to the M50 sqlite database is opened. The JTMS links associated with the M50, in each direction are also defined.

```{r}
direct <- "/var/datastore/2015/JTMS"
db <- src_sqlite("/var/datastore/SQLite/M50.sqlite", create = FALSE)

linksSB <- c(76, 61, 86, 17, 63, 58, 57, 91, 69, 53)
linksNB <- c(77, 66, 21, 19, 100, 52, 97, 70, 79, 94)
```

A function is created to iteratively read each JTMS csv file and load into a new table in the SQLite database.

```{r}
csv2SQLite <- function(path) {
    setwd(path)
    files <- list.files(pattern = "*.csv")
    names <- c("linkID", "UTC", "n", "outliers", "sum", "sumSquared")

    lapply(files, function(i) {
        a <- read_csv(i, col_names = names, skip = 1) %>%
            db_insert_into(con = db$con, table = "journeyTimes", values = .)
        
        rm(a)
        gc()
        })
}
```

In order to filter by UNIX timestamps in SQLite, a function is defined to list the timestamps of the start and end of each year between 2013 and 2016.

```{r}
createUnixCal <- function(year) {
    unixCal <- tibble(year = year, start = 0, end = 0) %>%
        mutate(start = as.POSIXct(paste(year, "-01-01 00:00:00", sep = "")),
               end = as.POSIXct(paste(year, "-12-31 23:59:59", sep = "")),
               startUnix = unclass(start),
               endUnix = unclass(end)) %>% 
        select(year, startUnix, endUnix)
}
```

Finally, a function is defined to split the journey times into seperate tables in the database for each year and for each direction. It also calculates the journey times in minutes

```{r}
filterJT <- function(year, links, direction = c("SB", "NB")) {
    
    unix <- createUnixCal(year)
    
    tbl(db, "journeyTimes") %>%
        filter(linkID %in% links) %>% 
        mutate(UTC = UTC / 1000) %>% 
        filter(UTC >= unix[[2]], UTC <= unix[[3]]) %>%
        collect(n = Inf) %>% 
        mutate(journeyTime = ifelse(n == 0, 0, ((sum/1000)/n)/60)) %>% 
        db_insert_into(con = db$con, table = paste("journeyTimes" , unix[[1]], 
                                                   direction, sep = "_"), values = .)
        }
```

Running the csv2SQLite function loads all JTMS data into the database

```{r, eval = FALSE}
csv2SQLite(direct)
```

```{r}
tbl(db, "journeyTimes") %>% 
    glimpse(width = 100)
```

Running the filterJT function splits the data into seperate tables.

```{r, eval = FALSE}
lapply((2013:2016), filterJT, linksSB, "SB")
lapply((2013:2016), filterJT, linksNB, "NB")
```

```{r}
tbl(db, "journeyTimes_2015_NB") %>% 
    glimpse(width = 100)
```

