---
title: "Joining Day Type data"
author: "Dan Brennan"
date: "10 March 2017"
output: github_document
---

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(purrr)
library(lubridate)
```

Connect to the M50 SQLite database

```{r}
db <- src_sqlite("/var/datastore/SQLite/M50.sqlite", create = FALSE)
```

Create a function to join day types to the C2 aggregated tables and JTMS tables

```{r}
joinDayTypes <- function(table) {
    dayTypes15 <- tbl(db, "dayTypes1219") %>% 
        collect(n = Inf) %>% 
        mutate(date = as.Date(as.POSIXct(date, origin = "1970-01-01"))) 
    
    lapply(table, function(row) {
         x <- tbl(db, as.character(row)) %>%
           collect(n = Inf)
         
         # c2 tables have "time" column, JTMS tables have "UTC column
         if ("time" %in% colnames(x)) {
           x <- x %>%
             mutate(time = as.POSIXct(time, origin = "1970-01-01"),
                    date = as.Date(time)) %>%
             left_join(dayTypes15, by = "date")
         }

         else {
           x <- x %>%
             mutate(time = as.POSIXct(UTC, origin = "1970-01-01"),
                    date = as.Date(time)) %>%
             left_join(dayTypes15, by = "date")
         }

         # # backup table
         saveRDS(x, paste("output/", as.character(row), ".rds", sep = ""))
         
         # drop from SQLite database and replace with updated table
         db$con %>%
           db_drop_table(as.character(row))
         
         db$con %>%
           db_insert_into(as.character(row), x)

         })
}

```

A dataframe of all C2 aggregated table names and JTMS table names is created.

```{r}
pattern <- "day_|hour_|min15_|min5_|journeyTimes_"

tables <- src_tbls(db) %>% 
  as_data_frame() %>% 
  filter(grepl(pattern, value)) %>% 
  as.list()

tables
```

The joinDayTypes function is then applied to all C2 aggregated tables and JTMS tables

```{r, eval = FALSE}
tables %>% 
  map(joinDayTypes)
```

Below is a sample of updated C2 hourly aggregated table

```{r}
tbl(db, "hour_50_2015") %>% 
    select(-date, -day) %>% 
    head(10)
```

Below is a sample of updated JTMS table

```{r}
tbl(db, "journeyTimes_2016_NB") %>% 
    select(-outliers, -time, -date, -day) %>% 
    head(10)
```

